---
title: "Statistics with R"
subtitle: "Introduction"
author: "Zhuanghua Shi (Strongway)"
date: "16 April 2018"
output: beamer_presentation
citation_package: natbib
link-citations: yes
header-includes:
 - \usetheme{metropolis}
 - \newcommand{\leftc}[1]{\begin{columns}\begin{column}{#1\textwidth}} 
 - \newcommand{\rightc}[1]{\end{column}\begin{column}{#1\textwidth}} 
 - \def\endc{\end{column} \end{columns}} 
 - \newcommand{\fig}[2]{\begin{figure} \centering \includegraphics[width=#1\textwidth]{images/#2}\end{figure}}
 - \newcommand{\figl}[2]{\begin{columns}\begin{column}{#1\textwidth}\begin{figure} \centering \includegraphics[width=.95\textwidth]{images/#2}\end{figure}}
 - \newcommand{\figr}[2]{\end{column}\begin{column}{#1\textwidth}\begin{figure} \centering \includegraphics[width=0.95\textwidth]{images/#2}\end{figure}\end{column} \end{columns}}
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
```

# Challenges in Research

\leftc{0.5}

* Psychological research requires
    * theory development
    * Hypothesis testing
* Empirical data analyses play key role in bridging the gap
    * Statistics knowledge
    * Statistics software (SPSS, Matlab, Excel, R)

\figr{0.5}{bridge.jpg}

# Why R?

\pause

* Open source
* Large community
* ...

__Most important features__

> R and RStudio constitute a powerful tool for reproducible research (RR)

# Reproducible Research (RR)

\leftc{0.5}

* Research process is not linear, rather often redo some steps. 
    * rerun experiments
    * reanalyze data

An ideal RR process should be able to adapt to flexible research process:

- easy to expand experiments and data analyses
- easy to maintain codes and text
- easy to replot figures and to output statistics 

\figr{0.5}{repeats.png}

# Solution: Git, R, and Markdown

- git for version controls (e.g., multiple experiments, minor variations)
- R for statistics
- RMarkdown for writing

__RStudio__ integrates all of them seaminglessly. This course we will focus on _R for Statistics_. 

# Aims of the course

\leftc{0.6}

* After this course, you should be able to  
    * Exploratory data analysis (EDA)
    * Import/Export various types of data
    * Basic statistics (ANOVA etc.)
    * Linear Regression
    * Multiple and logistic regression

* Lecturer and Tutors
    * Strongway (me)
    * Jan Nasenmann
    * Artyom Zinchenko
    
\figr{0.4}{goals.jpg}

# Schedule

```{r, comment=NA,message=FALSE, results='asis', warning = FALSE}
dates <- seq(as.Date('2018-04-16'), length.out = 12,by='1 week')
dates <- as.character(dates) # bug in xtable
contents <- c('Introduction', 
              'Data Visualiation',
              'Practice 1',
              '--- (Online assignment)',
              'Data Wrangling',
              '--- (Online assignment)',
              'Hypothesis Testing',
              'Practice 2',
              'Linear Regression',
              'Practice 3',
              'Multiple and Logistic Regression',
              'Practice 4'
              )
kable(data.frame(dates, contents))
```

# Online learning materials and sources

* The online book [An Introduction to Statistical and Data Sciences via R](http://www.moderndive.com)

This book will be our reference book of the course.

* [Datacamp](https://www.datacamp.com) 

We have a learning group in datacamp - [Statistics with R](https://www.datacamp.com/enterprise/statistics-with-r-5846c443-c1c8-4a98-9488-ed4290cc91d4). Please join in this group via the email I sent. Datacamp has huge learning materials, providing not only R but also Python and others. Please use this opportunity to learn data science. 

# Start with R and RStudio

* R, RStudio, and Packages
    * Please read the [chapter 2](http://www.moderndive.com/2-getting-started.html)

* Get familiar with RStudio Inferface

# Using R project

* R projects
    * make it straightforward to manage your files
    * are platform independent
    * have a root folder
    * Version control
    * .Rproj extension and one hiden folder
    * Store your history .Rhistory

Using R projects make you easy to run your codes across platforms and collaborate with colleagues. 

# Data Import/Export

- We encounter many types of data formats
    - csv, excel, matlab, spss, web table, Sql
- R {utils} provides basic import methods
    - read.csv
    - read.delim
    - read.table
- data.table package provides powerful function
    - fread()
- readr package (fast)
    - read_csv
    - read_delim
    - read_tsv
    
## import text file basic functions

> read.table(filename, header=TRUE, sep=',', stringsAsFactors = FALSE, col.names = c(...))

- specific functions:
    - read.csv()
    - read.csv2()
    - read.delim()
    - read.delim2()
- helper function
    - file.path()
    - past0()

## Comparison between utils and readr

utils          readr
------         ----------
read.table()   read_delim()
read.csv()     read_csv()
read.delim()   read_tsv()

## fread() from data.table

- super powerful, fast
- automatically manage column names
- infer column types and separators
- select or drop columns

## Import matlab file

Two options for importing matlab data

- convert data format to csv within Matlab
- using package 'R.matlab' 
    - readMat




